Generative AI for image reconstruction in Intensity Interferometry: a first attempt

### 2025-12-07 Phase_II/draft/Result.tex
before final

\section{Images Reconstructed by the GAN}
In this section, we begin by discussing phase retrieval using hyperparameters, as mentioned earlier, followed by an analysis in which multiple sources are trained simultaneously. The best performance for image reconstruction has been observed with a learning rate of \(2 \cdot 10^{-4}\), a kernel size of 5×5, and equal noise percentages (alpha/beta = 1) applied to both the original and generated images. A batch size of 1 is used, and equal training is provided to both the Discriminator and the Generator.

\subsection{Predicted Image from the Trained GAN}
Fig.~\ref{fig:GAN} demonstrates the success of the GAN in training a model to reconstruct the images of  fast-rotating stars using their Intensity Interferometry (II) observation . The GAN was trained on the training datasets for 100,000 steps and subsequently tested on various validation datasets to produce predicted images of the stars. The training was performed on a CPU using two nodes, each with 96 threads, and required approximately 20 hours. In Fig.~\ref{fig:GAN}, four combined images illustrate the GAN's performance in reconstructing the stars' shape, size, and brightness distribution using II.
\begin{itemize}
\item{The left panel shows the signals collected from six baselines, which serve as the input for the Generator during training.}
\item{The first middle panel displays the real image, or ground truth, which the Disciminator loss function uses to distinguish from the images generated by the Generator. During training, the GAN aims to keep minimum difference between these ground truth image and generated images.}
\item{The second middle panel presents the reconstructed, or predicted, image produced by the trained GAN, highlighting its success in image reconstruction.}
\item{The right panel shows the difference between the ground truth and the predicted image in the interferometric plane. The uniform background color of the frames suggests that the GAN model picks up white noise in the interferometric plane.}
\end{itemize}
The predicted images in Fig.~\ref{fig:GAN} yield encouraging results, accurately conveying visual information about the source's size, shape, and brightness distribution across its surface using only six baselines. However, further improvements can be achieved by increasing the number of telescopes to maximize coverage of the $(u, v)$ plane, making the existing and upcoming CTAO an ideal candidate for this approach.

\begin{figure*}
	\includegraphics[width=.32\linewidth]{fig/moments/mom0.png}\hfil
	\includegraphics[width=.32\linewidth]{fig/moments/mom1.png}\hfil
	\includegraphics[width=.32\linewidth]{fig/moments/mom2.png}\hfil
	\caption{This set of figures shows the comparison of monopole, x-centroid, and y-centroid for ground truth and predicted images generated by trained GAN.}
	\label{fig:cen}
\end{figure*}
\subsection{Evaluation of GAN using Moments}
The reconstructed images are visually compelling, demonstrating the GAN's effectiveness in using II to reconstruct images. However, visual assessment alone is insufficient; statistical evaluation is necessary to validate the results. To achieve this, we employ image moments as a statistical method. Image moments capture key properties of the reconstructed objects—such as shape, size, and intensity distribution—by quantifying features like position, orientation, and brightness distribution. By comparing the moments of the GAN-generated images to those of the ground truth, we can objectively assess the consistency and accuracy of the reconstruction. This approach provides a reliable framework for evaluating reconstruction quality, as image moments can reveal subtle differences in geometric and intensity properties that might not be apparent through visual inspection alone.

The raw moment $M_{ij}$ of an image is defined as \citep{hu1962visual}
\begin{equation}
	M_{ij} = \sum_{x} \sum_{y} x^i y^j I(x, y)
	\label{eqn:Mom}
\end{equation}
where $I(x,y)$ represents the intensity at pixel $(x,y)$. The zeroth order raw moment, or monopole, represents the total intensity of an image. It is computed by summing all pixel values across the image, yielding an overall intensity measure. In this context, analyzing the monopole provides the total flux of fast-rotating stars. According to Eq.~\eqref{eqn:Mom}, the monopole of an image is calculated as:
\begin{equation}
	M_{00} = \sum_{x} \sum_{y} I(x, y).
\end{equation}
The left figure of Fig.~\ref{fig:cen} displays the monopole values for 50 reconstructed images. The plot reveals a linear relationship between the monopole of the ground truth (real image) on the $x$-axis and that of the predicted (reconstructed) image on the $y$-axis, consistent across sources of varying shapes and sizes. This linearity confirms that the predicted images have an overall intensity (flux) that closely matches the ground truth. However, while the monopole effectively represents the total brightness, it does not provide information about the position, shape, size, or detailed brightness distribution of the fast-rotating stars. For these aspects, higher-order moments are necessary.

The center of mass of the sky image of any stellar object, is determined by its centroid, which provides the $x$ and $y$ coordinates representing the spatial position of the image. This centroid is computed using the first-order raw moments in conjunction with the monopole (the zeroth-order moment). The formulations for the centroid along the $x$ and $y$ directions are given by:
\begin{eqnarray}
&&x_c = \frac{M_{10}}{M_{00}} = \frac{\sum_{x,y} x \cdot I(x,y)}{\sum_{x,y} I(x,y)} \nonumber \\
&&y_c = \frac{M_{01}}{M_{00}} = \frac{\sum_{x,y} y \cdot I(x,y)}{\sum_{x,y} I(x,y)}
\end{eqnarray}
Here, \(I(x,y)\) represents the intensity at pixel \((x,y)\), \(M_{00}\) is the monopole (total intensity), and \(M_{10}\) and \(M_{01}\) are the first-order raw moments along the $x$ and $y$ axes respectively. This formulation accurately captures the spatial center of mass of the stellar object in the image.

The middle and right panel of Fig.~\ref{fig:cen} compare the centroids ($x_{c}$, $y_{c}$) of 50 predicted images with their corresponding ground truths respectively. The clustering of centroids within a specific scale range across all results indicates that the reconstructed images accurately represent the spatial location of the fast-rotating star relative to the ground truth.

\begin{figure*}
  \includegraphics[width=.32\linewidth]{fig/moments/mom3.png}\hfil
  \includegraphics[width=.32\linewidth]{fig/moments/mom4.png}\hfil
  \includegraphics[width=.32\linewidth]{fig/moments/mom5.png}\hfil
  \caption{The second-order central moments provide information about the size and shape of stellar objects.  Shown here are all the second-order central moments for ground truth and predicted images generated by the trained GAN. From left to right these are $\mu_{11},\mu_{20},\mu_{02}$.}
	\label{fig:struc}
\end{figure*}
\begin{figure*}
  \includegraphics[width=.49\linewidth]{fig/moments/mom6.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom7.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom8.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom9.png}\hfil
  \caption{Shown here are all the third-order central moments for
    ground truth and predicted images generated by the trained GAN.
    They represent the skewness of the brightness distributions.  The
    panels in reading order show
    $\mu_{30},\mu_{03},\mu_{21},\mu_{12}$.}
  \label{fig:moments}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Furthermore, these calculated centroids are instrumental in analyzing the shape, size, and brightness distribution of the stars using higher-order image moments. To this end, the central moment of an image is calculated according to:
\begin{equation}
	\mu_{pq} = \frac{1}{M_{00}}\sum_{x} \sum_{y} (x - x_c)^p (y - y_c)^q I(x, y).
\end{equation}
The sum of \(p\) and \(q\) defines the order of the central moment.

Fig.~\ref{fig:struc} presents the second-order central moments (\(\mu_{11}, \mu_{20}, \mu_{02}\)), which are used to study the structure of a fast-rotating star along the line of sight (as explained in the upcoming subsection). All three plots demonstrate a linear relationship in the second-order moments, similar to the monopole, thereby confirming the success of applying the GAN to reconstruct images with II.

The brightness distribution is characterized by the skewness of the image, which is quantified by calculating the third-order central moments (\(\mu_{30}, \mu_{03}, \mu_{21}, \mu_{12}\)). Fig.~\ref{fig:moments} presents all third-order moments for both the ground truth and the reconstructed image. The skewness along the $x$ and $y$ axes (\(\mu_{30}\) and \(\mu_{03}\)) appears acceptable, as shown in both upper panel of Fig.~\ref{fig:moments}, where a linear relationship exists between the ground truth and predicted images. However, the other higher-order moments (\(\mu_{21}\) and \(\mu_{12}\))—particularly \(\mu_{12}\), as depicted in both lower panel of Fig.~\ref{fig:moments}—do not align as well. This indicates that further improvement is possible and should be investigated.

>>>>

\section{Image Reconstruction: Results and Evaluation}

In this section, we examine the performance of the GAN Model whose architecture and choice of hyperparameters have been discussed above. We subject the trained Model to the task of phase retrieval and image reconstruction on the Testing Dataset.

\subsection{Visual Evaluation of Images Predicted by the Model}
Fig.~\ref{fig:GAN} demonstrates the success of the trained Model in reconstructing the images of a sample of the fictitious fast-rotators drawn from the Testing Set.

The four panels from left to right in each row of Fig.~\ref{fig:GAN} represent the following:
\begin{itemize}
\item{The left panel represents the sparse II pattern obtained by the simulated observation using the four IACTs illustrated in Fig.\ref{fig:teles}. As explained earlier, this image acts as part of the input, namely its ``condition'' part. This image acts as the condition that the sky-images of the star generated by the Generator must conform to.}

\item{The second panel from left displays the real image, or ground truth target image, which the Discriminator loss function uses to distinguish from the images generated by the Generator.}

\item{The third panel from left presents the reconstructed (or predicted) image corresponding to the ground truth (second panel) and generated by the trained Model. The similarity of these two images indicates the success of the Model in its stated objective of image reconstruction.}

\item{The right panel shows the residuals between the ground truth target image and the predicted image in the interferometric plane. The small values are indicative of the success of training the Model.}
\end{itemize}

The predicted images and the residuals presented in the third and the fourth columns (from left) of Fig.~\ref{fig:GAN} show encouraging results, conveying visual information about the source's size, shape, and brightness distribution across its surface fairly accurately. This has been achieved on the basis of the input provided by II observation using only six baselines. Arguably, further improvements can be achieved by increasing the number of telescopes to enhance the coverage of the $(u, v)$ plane. A closer examination of this proposition might be able to contribute to the design and instrumentation aspects in the existing and upcoming CTAO.

\begin{figure*}
	\includegraphics[width=.32\linewidth]{fig/moments/mom0.png}\hfil
	\includegraphics[width=.32\linewidth]{fig/moments/mom1.png}\hfil
	\includegraphics[width=.32\linewidth]{fig/moments/mom2.png}\hfil

\caption{This set of figures shows comparisons of the monopole, and the coordinates of the centroids ($x_c,\, y_c$) of the pixel brightness distribution of the ground truth images and the corresponding images predicted (generated) by the trained model. The monopole moment of the images represents the overall pixel brightness of the images. The values, as expected of a trained Model, not only lie in same range, but lie close to the diagonal and show an approximate equality. The centroids of both the ground truth images and the predicted images lie within small ranges of pixel values as should be expected of an ideally trained model. Please see the relevant text for detailed comments.}
\label{fig:cen}
\end{figure*}

\subsection{Evaluation of the Model using Moments}
The reconstructed images are visually compelling, demonstrating the Model's effectiveness in using II to reconstruct images. However, visual assessment alone is insufficient; statistical evaluation of the generated images in comparison with the ground truth images is necessary to validate the results. We present here the results of our calculation and comparison of the moments of distribution pixel brightness in the target ``Ground Truth'' images and the ``Predicted'' images.

Image moments capture key properties of the reconstructed objects, such as shape, size, and intensity distribution, by quantifying features like position, orientation, and brightness distribution. By comparing the moments of the Model-generated images with those of the ground truth, we can objectively assess the consistency and accuracy of the reconstruction. This approach provides a reliable framework for evaluating reconstruction quality, as image moments can reveal subtle differences in geometric and intensity properties that might not be apparent through visual inspection alone.

The raw moment $M_{ij}$ of an image is defined as \citep{hu1962visual}
\begin{equation}
	M_{ij} = \sum_{x} \sum_{y} x^i y^j I(x, y)
	\label{eqn:Mom}
\end{equation}
where $I(x,y)$ represents the intensity at pixel $(x,y)$. The zeroth order raw moment, or monopole, represents the total intensity of an image. It is computed by summing all pixel values across the image, yielding an overall intensity measure. In this context, analyzing the monopole provides the total pixel brightness of the images of the fictitious stars. According to Eq.~\eqref{eqn:Mom}, the monopole of an image is calculated as:
\begin{equation}
	M_{00} = \sum_{x} \sum_{y} I(x, y).
\end{equation}

An ideally trained Model should predict the overall pixel brightness of the generated images to be in the same range of values, if not equal to those of the ground truth images. The left panel of Fig.~\ref{fig:cen} displays the monopole values for 100 reconstructed images. The plot reveals a broad agreement in the overall intensity of the ground truth targets and the predicted (generated) images. However, a slightly higher brightness of reconstructed images in the intermediate range of brightness is evident. This could, perhaps, be mitigated by tuning the Batch Size hyperparameter.

While the monopole effectively represents the total brightness, it does not provide information about the position, shape, size, or detailed brightness distribution of the fast-rotating stars. For these aspects, higher-order moments are necessary.

The first order raw moments normalized by the respective monopole moments of the images given by 
\begin{eqnarray}
&&x_c = \frac{M_{10}}{M_{00}} = \frac{\sum_{x,y} x \cdot I(x,y)}{\sum_{x,y} I(x,y)} \nonumber \\
&&y_c = \frac{M_{01}}{M_{00}} = \frac{\sum_{x,y} y \cdot I(x,y)}{\sum_{x,y} I(x,y)}
\end{eqnarray}
represent the centroids $(x_c, y_c)$ of the pixel brightness distribution of the images. An ideally trained GAN model should produce the centroids of the target and predicted images within a small range of pixels. A larger range would imply a low fidelity in learning the brightness distribution of ground truth images. The middle and right panels of Fig.~\ref{fig:cen} compare the centroids ($x_{c}$, $y_{c}$) of 100 predicted images with their corresponding ground truths respectively. We notice that the coordinates of the centroids of ground truth images are spread across a range of 2 pixels whereas the those of the predicted images are spread across a span of 4 pixels along $x$-axis and 7 pixels along $y$-axis. This clustering of centroids within a small scale range across all results indicates that the reconstructed images represent the spatial distribution of brightness of predicted and ground truth images fairly accurately.

\begin{figure*}
  \includegraphics[width=.32\linewidth]{fig/moments/mom3.png}\hfil
  \includegraphics[width=.32\linewidth]{fig/moments/mom4.png}\hfil
  \includegraphics[width=.32\linewidth]{fig/moments/mom5.png}\hfil
  \caption{The second order central moment of the brightness distribution of the images is analogous to the moment of inertia of a mass distribution. The panels of this figure show the comparison of the second order moments ($\mu_{11}$, $\mu_{20}$, and $\mu_{02}$) of the predicted images vs. the ground truth images from left to right respectively. Approximate equality of these moments is evident. The small scatter in the moments of the predicted images is indicative of a balanced training of the Model. }
\label{fig:struc}
\end{figure*}
\begin{figure*}
  \includegraphics[width=.49\linewidth]{fig/moments/mom6.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom7.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom8.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom9.png}\hfil
  \caption{Shown here are all the third-order central moments ($\mu_{30}$, $\mu_{03}$, $\mu_{21}$, and $\mu_{12}$) for ground truth and predicted images generated by the trained Model. They represent the skewness of the brightness distributions. The low skewness in the brightness distribution of the ground truth images is evident. One outlier among the predicted images dominates their skewness distribution along the $y$-axis as seen in the values of \(\mu_{03}\) and \(\mu_{12}\). The trained Model, however, seems to have learned the approximately uniform distribution of skewness along the $x$-axis as is evident, with a few outliers though, in the values of \(\mu_{30}\) and \(\mu_{21}\).}
\label{fig:moments}
\end{figure*}

Furthermore, these calculated centroids are instrumental in analyzing the shape, size, and brightness distribution of the stars using higher-order image moments. To this end, the central moment of an image is calculated according to:
\begin{equation}
	\mu_{pq} = \frac{1}{M_{00}}\sum_{x} \sum_{y} (x - x_c)^p (y - y_c)^q I(x, y).
\end{equation}
The sum of \(p\) and \(q\) defines the order of the central moment.

The second order central moment of the brightness distribution of the images is analogous to the moment of inertia of a mass distribution. Fig.~\ref{fig:struc} presents the comparison of second-order central moments (\(\mu_{11}, \mu_{20}, \mu_{02}\)), which are used to study the structure of a fast-rotating star along the line of sight (as explained in the upcoming subsection). All three plots demonstrate an approximate equality among these moments, similar to the monopole, thereby confirming the success of applying the Model to reconstruct images with II. The small scatter in the moments of the predicted images indicates robust learning of the Model without over-fitting.

The third-order central moments (\(\mu_{30}, \mu_{03}, \mu_{21}, \mu_{12}\)) of the images quantify the degree and direction of asymmetry of the brightness distribution around their respective centroids. Fig.~\ref{fig:moments} presents a comparison of all third-order moments for both the ground truth and the reconstructed image.    All the four panels show a low skewness in the brightness distribution of the ground truth images. The skewness of the predicted images along the $y$-axis is dominated by one outlier as seen in the values of \(\mu_{03}\) and \(\mu_{12}\). However, the approximately uniform distribution of skewness of the predicted images along the $x$-axis, with a few outliers though, as seen in the values of \(\mu_{30}\) and \(\mu_{21}\) suggests that this feature in the ground truth images has been picked up fairly well during the training of the Model.

### 2025-12-07 Phase_II/draft/conclusion.tex
before final

This work addresses the challenge of phase retrieval in II using a machine-learning technique, specifically a conditional Generative Adversarial Network (cGAN). Our study demonstrates \st{that applying a cGAN to II data successfully recovers the size, shape, and brightness distribution of a fast-rotating star.}  \textcolor{blue}{that the size, shape and brightness distribution of fast rotating stars can be recovered by a cGAN trained on the combined input of the sky-image of known sources  along with their respective II data. In this training the sky-image acts as the real ``ground truth'' and the II data acts as the ``condition''. The Discriminator of the cGAN model is trained to efficiently distinguish between the real images and fake (generated) images based on the ``ground truth'' images and the respective II data condition. The Generator is trained to generate progressively realistic images which are also consistent with the condition of the II data. The evaluation of the trained model is then carried out} \st{Evaluations based on} \textcolor{blue}{by comparison of image moments of real images and generated images.} Specifically, the monopole, second, and third-order moments \textcolor{blue}{are compared. The results} support the effectiveness of cGAN in achieving \textcolor{blue}{the stated objective.} \st{accurate image reconstruction from a simulation of II from a single site with four telescopes.}

While the results of this study highlight the significant potential of machine learning, and in particular the applicability of cGAN, for image reconstruction in II, several aspects require further refinement. First, an important factor in the reconstruction process is the extent of Fourier plane coverage, which depends on the number of available telescopes and the total observing time. The reasonable success of this piece of work suggests that a network of higher number of telescopes providing higher number of baselines and greater coverage of the $(u,v)$ plane signal, \textcolor{blue}{would play a critical role in} projects of image reconstruction of more complicated stellar systems can be undertaken. Future work might explore different observatory layouts to assess their impact on image reconstruction quality.  Second, detector efficiencies, which impact the signal-to-noise ratio (SNR) of actual observational data, have not yet been incorporated; addressing these factors will be crucial for more accurate SNR estimation.  Third, exploring and comparing alternative methods for image generation could reveal approaches that outperform cGAN in reconstructing stellar images with II. Fourth, experimenting with different loss functions could provide additional insights into the reconstruction quality. Although further testing is needed to refine the GAN and enhance its robustness and reliability, our findings suggest that machine learning is a promising approach for phase reconstruction in II.

>>>>

This work addresses the challenge of phase retrieval in II using a machine-learning technique, specifically a conditional Generative Adversarial Network (cGAN). Our study demonstrates that the size, shape and brightness distribution of fast rotating stars can be recovered by a Pix2Pix cGAN model trained on the combined input of the sky-image of known sources  along with their respective II data. In this training the sky-image acts as the real ``ground truth'' and the II data acts as the ``condition''. The Discriminator of our Model is trained to efficiently distinguish between the real images and fake (generated) images based on the ``ground truth'' images and the respective II data as the condition. The Generator is trained to generate progressively realistic images which are also consistent with the condition of the II data. The evaluation of the trained Model is then carried out by comparison of image moments of ground truth images and generated images. Specifically, the monopole, second, and third-order moments are compared. The results support the effectiveness of cGAN in achieving the stated objective.

While the results of this study highlight the significant potential of machine learning, and in particular the applicability of cGAN, for image reconstruction in II, several aspects require further refinement. First, an important factor in the reconstruction process is the extent of Fourier plane coverage, which depends on the number of available telescopes and the total observing time. The reasonable success of this piece of work suggests that a network of higher number of telescopes providing higher number of baselines and greater coverage of the $(u,v)$ plane signal, would play a critical role in projects of image reconstruction of more complicated stellar systems can be undertaken. Future work might explore different observatory layouts to assess their impact on image reconstruction quality.  Second, detector efficiencies, which impact the signal-to-noise ratio (SNR) of actual observational data, have not yet been incorporated; addressing these factors will be crucial for more accurate SNR estimation.  Third, exploring and comparing alternative methods for image generation could reveal approaches that outperform cGAN in reconstructing stellar images with II. Fourth, experimenting with different loss functions could provide additional insights into the reconstruction quality. Although further testing is needed to refine the GAN and enhance its robustness and reliability, our findings suggest that machine learning is a promising approach for phase reconstruction in II.

### 2025-12-07 Phase_II/draft/image-reconstruction-past-efforts.tex
Edit of GAN drat completed by SS
before final

\section{Past Efforts at Image Reconstruction in Intensity Interferometry}
\iffalse
Several theoretical and computational approaches for phase reconstruction with II have been proposed. \cite{gamo1963triple} introduced the concept of triple-intensity correlation, which \cite{goldberger1963use} subsequently applied in an experiment to observe scattered particles in microscopic systems. Sato conducted experiments to measure the diameter and phase of asymmetrical objects, suggesting that triple correlation could extend II to image stellar bodies \citep{sato1978imaging, sato1979computer, sato1981adaptive}. However, achieving a satisfactory signal-to-noise ratio (SNR) remained a significant challenge for this approach.

\cite{GerchbergSaxton1972} suggested an iterative method to determine the phase from the image and diffraction plane pictures. This method relies on accurate initial estimates and is vulnerable to slow convergence otherwise. \cite{Fienup1982} introduced a Hybrid Input-Output algorithm that incorporates feedback mechanisms to improve convergence rate and robustness, particularly in noisy environments.

Later, \cite{holmes2010two} proposed an alternative method that utilizes the Cauchy-Riemann relations to reconstruct 1-D images. They also extended the approach to 2-D images across a range of signal-to-noise (SNR) values. This algorithm was applied to simulated data of stellar objects using II, considering both existing and forthcoming Imaging Cherenkov Telescope Arrays (IACTs) with a large number of telescopes \citep{nunez2010stellar, nunez2012high, nunez2012imaging}. However, this method faces challenges related to computational complexity when attempting to generalize to higher dimensions.

\cite{Li2014} suggested a flexible iterative Regularization method that incorporates prior information (e.g., sparsity, smoothness, or non-negativity) to reduce the ill-posedness of the phase retrieval problem. This method is more robust against noise and stabilizes the solution against artefacts and spurious solutions. Nevertheless, it faces challenges regarding the choice of the regularization parameter, computational complexity, and sensitivity to the initial guess.

The Transport-of-Intensity Equation (TIE) method is a non-interferometric technique first proposed by \cite{Teague1983} that relates the intensity variations along the optical axis to the phase of the optical fields. This method enables phase retrieval from intensity measurements taken at multiple planes. \cite{Zhang2020} proposed a method to obtain a \textquotedblleft universal solution\textquotedblright\ to the TIE by employing a \textquotedblleft maximum intensity assumption\textquotedblright, thereby converting the TIE into a Poisson equation which is then solved iteratively. More recently, \cite{Kirisits2024} have explored hybrid methods that combine the TIE with other equations, such as the Transport of Phase Equation (TPE). These approaches leverage the strengths of both equations to improve phase retrieval accuracy. This method is universally applicable, as it works for arbitrarily shaped apertures, handles non-uniform illumination, and accommodates inhomogeneous boundary conditions. It guarantees convergence, although the speed of convergence depends on the quality of the initial guess, and the final results are influenced by the boundary conditions.

%%%**************************8

%\documentclass{article}
%\usepackage{amsmath}
%\begin{document}
\fi
Several approaches have been developed for phase reconstruction in intensity interferometry (II). \cite{gamo1963triple} introduced triple-intensity correlation, applied by \cite{goldberger1963use} to microscopic systems and extended by \cite{sato1978imaging, sato1979computer, sato1981adaptive} to measure stellar diameters and phases, though limited by low signal-to-noise ratio (SNR). \cite{GerchbergSaxton1972} proposed an iterative phase retrieval method using image and diffraction plane data, sensitive to initial estimates and convergence speed. \cite{Fienup1982} improved this with the Hybrid Input-Output algorithm, enhancing robustness in noisy conditions. \cite{holmes2010two} utilized Cauchy-Riemann relations for 1-D and 2-D image reconstruction, applied to simulated stellar data with Imaging Cherenkov Telescope Arrays \citep{nunez2010stellar, nunez2012high, nunez2012imaging}, but faced computational complexity for higher dimensions. \cite{Li2014} introduced a regularized iterative method incorporating priors (e.g., sparsity) to mitigate noise and ill-posedness, though challenged by parameter tuning and initial guess sensitivity. The Transport-of-Intensity Equation (TIE), proposed by \cite{Teague1983}, retrieves phase from intensity variations across planes; \cite{Zhang2020} solved TIE as a Poisson equation using a maximum intensity assumption, while \cite{Kirisits2024} combined TIE with the Transport of Phase Equation for improved accuracy across arbitrary apertures and non-uniform illumination, with convergence dependent on initial guesses and boundary conditions.

%\end{document}
%%%**************

@@

In the astrophysical context, \cite{schawinski2017galaxypics} employed a GAN model to recover features — such as spiral arms, central bulges, and disk structures of galaxies — from noise-affected images. \cite{mustafa2019cosmogan} developed and customized a Deep Convolutional GAN, dubbed \textquotedblleft CosmoGAN\textquotedblright, capable of generating high-fidelity weak-lensing convergence maps of dark matter distribution that statistically reproduce real weak lensing structures. \cite{coccomini2021lightweightgan} have successfully generated credible images of planets, nebulae, and galaxies using \textquotedblleft lightweight\textquotedblright\ and \textquotedblleft physics-uninformed\textquotedblright\ GANs to produce synthetic images of celestial bodies. They also generated a \textquotedblleft Hubble Deep Field-inspired\textquotedblright\ wide-view simulation of the universe.

Let us now discuss the details of our procedure and method.

>>>>

\section{Past Efforts at Image Reconstruction in Intensity Interferometry}
Several approaches have been developed for phase reconstruction in intensity interferometry (II). \cite{gamo1963triple} introduced triple-intensity correlation, applied by \cite{goldberger1963use} to microscopic systems and extended by \cite{sato1978imaging, sato1979computer, sato1981adaptive} to measure stellar diameters and phases, though limited by low signal-to-noise ratio (SNR). \cite{GerchbergSaxton1972} proposed an iterative phase retrieval method using image and diffraction plane data, sensitive to initial estimates and convergence speed. \cite{Fienup1982} improved this with the Hybrid Input-Output algorithm, enhancing robustness in noisy conditions. \cite{holmes2010two} utilized Cauchy-Riemann relations for 1-D and 2-D image reconstruction, applied to simulated stellar data with Imaging Cherenkov Telescope Arrays \citep{nunez2010stellar, nunez2012high, nunez2012imaging}, but faced computational complexity for higher dimensions. \cite{Li2014} introduced a regularized iterative method incorporating priors (e.g., sparsity) to mitigate noise and ill-posedness, though challenged by parameter tuning and initial guess sensitivity. The Transport-of-Intensity Equation (TIE), proposed by \cite{Teague1983}, retrieves phase from intensity variations across planes; \cite{Zhang2020} solved TIE as a Poisson equation using a maximum intensity assumption, while \cite{Kirisits2024} combined TIE with the Transport of Phase Equation for improved accuracy across arbitrary apertures and non-uniform illumination, with convergence dependent on initial guesses and boundary conditions.

@@

In the astrophysical context, \cite{schawinski2017galaxypics} employed a GAN model to recover features, such as spiral arms, central bulges, and disk structures of galaxies, from noise-affected images. \cite{mustafa2019cosmogan} developed and customized a Deep Convolutional GAN, dubbed \textquotedblleft CosmoGAN\textquotedblright, capable of generating high-fidelity weak-lensing convergence maps of dark matter distribution that statistically reproduce real weak lensing structures. \cite{coccomini2021lightweightgan} have successfully generated credible images of planets, nebulae, and galaxies using \textquotedblleft lightweight\textquotedblright\ and \textquotedblleft physics-uninformed\textquotedblright\ GANs to produce synthetic images of celestial bodies. They also generated a \textquotedblleft Hubble Deep Field-inspired\textquotedblright\ wide-view simulation of the universe.

### 2025-12-07 Phase_II/draft/intro_reformatted.tex
before final

This paper is organized as follows. The next section discusses briefly the past efforts at image reconstruction on II, followed by the section on a discussion of Intensity Interferometry, focusing on its signal and noise characteristics for fast-rotating stars along the Earth’s rotation. The following section introduces the cGAN formulation and its structure. The fifth section details the parameter selection for training the network for image reconstruction. The sixth section presents the results of the trained network both visually and via image moments. Finally, the paper concludes with a discussion of the overall results.
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{fig/telescope.png}
	\caption{A schematic representation of {\textcolor{blue}{a hypothetical observation facility with an the array of four Cherenkov Telescopes (IACTs) whose relative positions approximately mimic those at VERITAS. This array is used in simulating the Intensity Interferometry (II) observation of the fast rotators.}} Each of the telescopes has a diameter of 12m. The baselines provided by the array are of the order of 100m.}
	\label{fig:teles}
\end{figure}

>>>>

This paper is organized as follows. The next section discusses briefly the past efforts at image reconstruction on II, followed by the section on a discussion of II, focusing on its signal and noise characteristics for fast-rotating stars along the Earth’s rotation. The following section introduces the cGAN formulation and its structure. The fifth section details the parameter selection for training the network for image reconstruction. The sixth section presents the results of the trained network both visually and via image moments. Finally, the paper concludes with a discussion of the overall results.
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{fig/telescope.png}
	\caption{A schematic representation of a hypothetical observation facility with an the array of four Cherenkov Telescopes (IACTs) whose relative positions approximately mimic those at VERITAS. This array is used in simulating the II observation of the fast rotators. Each of the telescopes has a diameter of 12m. The baselines provided by the array are of the order of 100m.}
	\label{fig:teles}
\end{figure}

### 2025-12-07 Phase_II/draft/main.tex
before final

\documentclass[linenumbers, twocolumn]{aastex701}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
%\newcommand{\paperdoi}{10.5281/zenodo.17598807}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyfoot{}
\fancyfoot[L]{DOI: \href{https://doi.org/10.5281/zenodo.17598807}{10.5281/zenodo.17598807}}

@@

%\maketitle
%\begin{center}
%\textbf{DOI:} \href{https://doi.org/\paperdoi}{\paperdoi}
%\end{center}

\begin{abstract}
In the last few years Intensity Interferometry (II) has made significant strides in achieving high-precision resolution of stellar objects at optical wavelengths. Despite these advancements, phase retrieval remains a major challenge due to the nature of photon correlation. This paper explores the application of a conditional Generative Adversarial Network (cGAN) to tackle the problem of image reconstruction in II. This approach successfully reconstructs the shape, size, and brightness distribution of a fast-rotating star from sparsely sampled, spatial power spectrum of the source, corresponding to II with four telescopes.  Although this particular example could also be addressed using parameter fitting, the results suggest that with larger arrays much more complicated systems could be reconstructed by applying machine-learning techniques to II.
\end{abstract}

\input{intro_reformatted}
\input{image-reconstruction-past-efforts}
\input{II}
\input{GAN}
\input{ParaGan}
\input{Result}
\input{conclusion}
\begin{acknowledgements}
\textbf{Acknowledgment:} One of the authors (SS) gratefully acknowledges the computing facilities and the local hospitality extended to him by the Inter-University Centre for Astronomy and Astrophysics (IUCAA), Pune, India under its Visiting Associate Programme during the preparation of this manuscript.
\end{acknowledgements}

>>>>

\documentclass[linenumbers, twocolumn]{aastex701}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage[normalem]{ulem}
\usepackage{hyperref}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyfoot[L]{DOI: \href{https://doi.org/10.5281/zenodo.17598807}{10.5281/zenodo.17598807}}

@@

\begin{abstract}
In the last few years Intensity Interferometry (II) has made significant strides in achieving high-precision resolution of stellar objects at optical wavelengths. Despite these advancements, phase retrieval remains a major challenge due to the nature of photon correlation. This paper explores the application of a conditional Generative Adversarial Network (cGAN) to tackle the problem of image reconstruction in II. This method successfully reconstructs the shape, size, and brightness distribution of simulated, fast-rotating stars based on a sparsely sampled spatial power spectrum obtained from a hypothetical ground-based II facility composed of four Imaging Atmospheric Cherenkov Telescopes (IACTs).  Although this particular example could also be addressed using parameter fitting, the results suggest that with larger arrays much more complicated systems could be reconstructed by applying machine-learning techniques to II.
\end{abstract}

\input{intro_reformatted}
\input{image-reconstruction-past-efforts}
\input{II}
\input{GAN}
\input{ParaGan}
\input{Result}
\input{conclusion}
\begin{acknowledgements}
\textbf{Acknowledgment:} One of the authors (SS) gratefully acknowledges the computing facilities and the local hospitality extended to him by the Inter-University Centre for Astronomy and Astrophysics (IUCAA), Pune, India under its Visiting Associate Programme during the preparation and finalization of this manuscript.
\end{acknowledgements}
